---
title: 'Plots: Graphical Methods of Data Exploration'
output:
html_document:
df_print: paged
---

A scatterplot is the most appropriate plot if the data being used is best considered to be continuous for the variable in question. 

Allocation involves initializing an object to be of any type and value. 

```{r}
x= 1:10
x

```
Concatenation is the joining of values, in this case, into one vector. This is done using the c() function.

```{r}
y=c(11,12,9,7,5,8,4,4,5,3)
y

```

Plot directive needs two arguments: one for x-axis, another for y-axis
Visualize the points from x, y

```{r}
plot(x,y)

```
Add labels to axes 

```{r}
plot(x,y,ylab = "Response variable",xlab = "Explanatory variable")

```
Change plotting symbols.

```{r}
plot(x,y,pch=3,ylab = "Response variable",xlab = "Explanatory variable")
```

```{r}
plot(x,y,pch=2,ylab = "Response variable",xlab = "Explanatory variable")
```
Add lines to scatterplot to for:
1. Determining regression
2. Own specifications

```{r}
plot(x,y,pch=2,ylab="Response variable",xlab="Explanatory variable",lines(c(0,10),c(12,0),lty=2))
abline(lm(y~x))
```


```{r}
v = c(2,4,6,8,10)
w = c(8,5,6,6,2)
```

```{r}
plot(x,y,pch=2,ylab="Response variable",xlab="Explanatory variable")
abline(lm(y~x))
lines(c(0,10),c(12,0),lty=2)
points(v,w,pch=3)
abline(lm(w~v),lty=3)
```
Use dataframe called "weather" and read data from the file to use
```{r}
sex=c("male","female")
weather<-read.table("/Users/mattjchin/Desktop/personal_projects/r_work/Course Downloadable Data  Exercises-20220120/SilwoodWeather.txt",header=T)
names(weather)
```
Attach the names info to the data frame

```{r}
attach(weather)
month = factor(month)

# Check status of variable "month"
is.factor(month)
```
Variable status of factor confirmed. 
Plot using categorical explanatory variable, in this case, month
```{r}
plot(month,upper)
```
Data Visualization in Color
This is a color wheel of 30 colors
```{r}
pie(rep(1,30), col=rainbow(30), radius=0.9)
```
Creating a color wheel with 10 colors and radius 0.5
Colors created are generated in counterclockwise direction
```{r}
pie(rep(1,10), col=rainbow(10), radius=0.5)
```

```{r}
x = seq(0,10,0.1)
y1 = 2 + 3 * x- 0.25*x^2
y2 = 3+3.3*x- 0.3 * x^2
```

```{r}
par(bg="ghostwhite")
plot(x,y2,type="n",ylab = "")
lines(x,y2,col="red")
lines(x,y1,col="blue")
```
Colored scatterplot

```{r}
jantemps=read.table("/Users/mattjchin/Desktop/personal_projects/r_work/Course Downloadable Data  Exercises-20220120/jantemp.txt", header=T)
attach(jantemps)
names(jantemps)
```

```{r}
max(tmax)
```

```{r}
min(tmin)
```
Includes for loop in R
```{r}
plot(day,tmax,ylim=c(-12,12),type="n",ylab = "Temperature")
points(day,tmin,col="blue",pch=16)
points(day,tmax,col="red",pch=16)
for (i in 1:31) lines(c(i,i), c(tmin[i], tmax[i]), col="green")
```
Histogram
```{r}
x = rnorm(1000)
par(bg="cornsilk")
hist(x, col="lavender", main="")
```
Pie chart with color
```{r}
fate = c(0.12, 0.3, 0.26, 0.16, 0.04, 0.12)
names(fate) = c("Ragwort", "Thistle", "Willowherb", "Rush", "Orchid", "Knapweed")
pie(fate, col = c("purple", "violetred1", "green3", "cornsilk", "cyan", "white"))
```
Multivariate Plots
```{r}
pollution = read.table("/Users/mattjchin/Desktop/personal_projects/r_work/Course Downloadable Data  Exercises-20220120/OtherFiles/pollute.txt",header=T)
attach(pollution)
names(pollution)
 
```
Scatterplots, trendlines all in an entire matrix
```{r}
pairs(pollution,panel=panel.smooth)
```
Tree-based models
```{r}
library(tree)
regtree = tree(Pollution~., data=pollution)
```

```{r}
plot(regtree)
text(regtree)
```

```{r}
attach(pollution)
coplot(Pollution~Temp | Rain)
```

```{r}

par(mfrow=c(1,1))
par(mfrow=c(1,2))
par(mfrow=c(2,2))

```

```{r}
plotdata = read.table("/Users/mattjchin/Desktop/personal_projects/r_work/Course Downloadable Data  Exercises-20220120/plotdata.txt", header = T)
attach(plotdata)
names(plotdata)

```

```{r}
par(mfrow=c(2,2))
plot(x,y,type="l")
plot(x,y,log="xy",type="l")
plot(x,y,log="y", type="l")
plot(x,y,log="x", type="l")
```

```{r}
par(mfrow=c(1,2))
plot(x,y,type="l")
plot(x,y,ylim=c(0,50),type="l")
text(0.8,45,"(b)")
```

```{r}
map.places= read.csv("/Users/mattjchin/Desktop/personal_projects/r_work/Course Downloadable Data  Exercises-20220120/map.places.csv.txt",header=T)
attach(map.places)
names(map.places)
```

```{r}
map.data=read.csv("/Users/mattjchin/Desktop/personal_projects/r_work/Course Downloadable Data  Exercises-20220120/OtherFiles/bowens.csv",header=T)
attach(map.data)
names(map.data)
nn=ifelse(north<60,north+100,north)
```

```{r}
plot(c(20,100),c(60,110),type="n",xlab="",ylab="")
for (i in 1:length(wanted)){
ii <- which(place == as.character(wanted[i]))
text(east[ii], nn[ii], as.character(place[ii]), cex = 0.6) }

```

```{r}
```

```{r}
labels=letters[1:10]
labels
```

```{r}
plot(1:10,1:10,type="n")
text(1:10,1:10,labels,cex=2)
plot(1:10,1:10,type="n")
text(1:10,10:1,labels,cex=2,srt=180)
```








