---
title: "Matthew Chin - Mathematical Equations in R"
output:
  html_document:
    df_print: paged
---


```{r}
# rm(x) # There is no variable named "x", commented out
# Read data from the mode text file
distribution = read.table("mode.txt", header = T)
attach(distribution)
names(distribution)
```

Show the values in distribution table
```{r}
distribution
```


Mode
```{r}
par(mfrow=c(1,2))
barplot(fx, names=as.character(x))
barplot(fy,names=as.character((x)))
```

Median
```{r}
y = rep(x, fx)
y = sort(y)
y
```

```{r}
length(y)
```

```{r}
ceiling(length(y)/2)
```
```{r}
y[17]
```
Median
```{r}
median(y)
```
Arithmetic mean
```{r}
sum(y)/length(y)
```
Mean
```{r}
mean(y)
```
Geometric mean
```{r}
prod(y)^(1/length(y))
```
Mean of log(y)
```{r}
meanlogy = sum(log(y))/length(y)
meanlogy
```

```{r}
exp(meanlogy)
```
Anti-log of mean(log(y))
```{r}
geometric = function(x) exp(sum(log(x))/length(x))
geometric(y)
```
Get mean of aphid
```{r}
aphid = c(10, 1, 1, 10, 1000)
mean(aphid)
```
Get geometric mean of aphid
```{r}
geometric(aphid)
```
Harmonic mean
```{r}
mean(c(1,2,4,1))
```

```{r}
4/sum(1/c(1,2,4,1))
```

```{r}
harmonic = function(x) 1/mean(1/x)
harmonic(y)

```

```{r}
mode(y)
```

```{r}
par(mfrow=c(1,1))
barplot(fz, names=as.character(x))
```

```{r}
w = rep(x,fz)
w
```
Measuring variation
```{r}
y = c(13,7,5,12,9,15,6,11,9,7,12)
plot(y,ylim = c(0,20))

```

```{r}
range(y)
```

```{r}
plot(y,ylim = c(0,20))
abline(mean(y), 0)
```

```{r}
A=c(3,4,4,3,2,3,1,3,5,2)
B=c(5,5,6,7,4,4,3,5,6,5)
C=c(3,3,2,1,10,4,3,11,3,10)
```

```{r}
mean(A)
```

```{r}
mean(B)
```

```{r}
mean(C)
```

```{r}
dA = A-3
dA
```

```{r}
dB = B-5
dB
```

```{r}
dC = C-5
dC
```


```{r}
SSA = sum(dA^2)
SSB = sum(dB^2)
SSC = sum(dC^2)
```


```{r}
SSA;SSB;SSC
```

```{r}
s2A = SSA/9
s2B = SSB/9
s2C = SSC/9
```


```{r}
s2A;s2B;s2C
```

```{r}
s2A = var(A)
s2A
```

```{r}
14.2222/1.3333
```

```{r}
sqrt(s2A/10)
```

```{r}
sqrt(s2B/10)
```

```{r}
sqrt(s2C/10)
```

```{r}
qt(.025,9)
```

```{r}
qt(.975,9)
```

```{r}
qt(.995,9)
```

```{r}
qt(.9975,9)
```

```{r}
qt(.975,9)*sqrt(1.33333/10)
```

```{r}
z=rnorm(1000)
# z
mean(z)
```

```{r}
quantile(z, c(.025, .975))
```

```{r}
z<-rnorm(10000)
quantile(z,c(.025,.975))
```

```{r}
y = c(3,4,6,4,5,2,4,5,1,5,4,6)
mad(y)
  
```

```{r}
sd(y)
```

```{r}
y1 = c(y, 100)
mean(y1)
```

```{r}
sqrt(var(y1))
```

```{r}
mad(y1)
```
Outliers
```{r}
outlier = function(x) {
  if(sqrt(var(x))>4*mad(x)) print("Outliers present")
  else print("Deviation reasonable")}

outlier(y)
```

```{r}
outlier(y1)
```
Single-sample estimation
```{r}
light = read.table("light.txt", header = T)
attach(light)
light
```

```{r}
names(light)
```

```{r}
hist(speed)
```
Get summary of speed
```{r}
summary(speed)
```

```{r}

wilcox.test(speed,mu=990)
```
Thus, we accept the alternative hypothesis since p=.00213<.05 for a 95% confidence interval. 

Student t-test
```{r}
t.test(speed, mu=990)
```

```{r}
qt(.975,18)
```

```{r}
(mean(A)-mean(B))/sqrt(s2A/10+s2B/10)
```

```{r}
t.test(A,B)
```
Wilcoxon rank sum test
```{r}
par(mfrow=c(1,2))
hist(B,breaks=c(0.5:11.5))
hist(C,breaks=c(0.5:11.5))
```

```{r}
combined = c(B,C)
combined
```

```{r}
sample = c(rep("B",10),rep("C",10))
sample
```

```{r}
rank.combi = rank(combined)
rank.combi
```

```{r}
sum(rank.combi[sample=="B"])
```

```{r}
sum(sum(rank.combi[sample=="C"]))
```

```{r}
tapply(rank.combi, sample, sum)
```

```{r}
wilcox.test(B,C)
```

```{r}
x=c(20,15,10,5,20,15,10,5,20,15,10,5,20,15,10,5)
y=c(23,16,10,4,22,15,12,7,21,16,11,5,22,14,10,6)
```

```{r}
t.test(x,y)
```

```{r}
t.test(x,y,paired=T)
```

```{r}
par(mfrow=c(1,2))
y=rnbinom(1000,1,0.2)
mean(y)
```

```{r}
var(y)
```

```{r}
table(y)
```

```{r}
hist(y,breaks=-0.5:38.5)
```


```{r}
my = numeric(1000)
for (i in 1:1000) {
  y = rnbinom(30,1,0.2)
  my[i] = mean(y)}
hist(my)
```

```{r}
par(mfrow=c(1,1))
```






