---
title: "Matthew Chin - Project 4 in R"
output:
  pdf_document: default
  html_notebook: default
---

 
1. Read the data into data frame
```{r}
# Copied data from Excel file and pasted into and saved as CSV
cgdata = read.csv("CollegeGrads.csv", header = TRUE)
cgdata

```

2. Check the column headings
```{r}
names(cgdata)
```


3. Check data type, structure, and search for missing data
```{r}
typeof(cgdata)
typeof("State")
typeof("ColGrad.")
typeof("Dropout.")
typeof("EdSpend")
typeof("Metro.")
typeof("Age")
typeof("LPRFem")
typeof("Neast")
typeof("Seast")
typeof("West")
typeof("Midwest")
```


```{r}
summary(cgdata)
```


```{r}
head(cgdata)
```

Return only first 10 states' data and first 5 columns
```{r}
cgdata[1:10, 1:5]
```

Get the structure of the data frame
```{r}
structure(cgdata)
```

Check for null (NA) values in column one
```{r}
is.na(cgdata[1])
```

Find if any of the other nine columns have any missing NA data, using which() function
```{r}
which(is.na(cgdata[2]))
which(is.na(cgdata[3]))
which(is.na(cgdata[4]))
which(is.na(cgdata[5]))
which(is.na(cgdata[6]))
which(is.na(cgdata[7]))
which(is.na(cgdata[8]))
which(is.na(cgdata[9]))
which(is.na(cgdata[10]))
```

```{r}
cgdata[1,2]
```

```{r}
st_labs = "AL AK AZ AR CA CO CT DE FL GA HI ID IL IN IA KS KY LA ME MD MA MI MN MS MO MT NE NV NH NJ NM NY NC ND OH OK OR PA RI SC SD TN TX UT VT VA WA WV WI WY"
st_labs
```


```{r}
barplot(cgdata$EdSpend,main="Education Spending by State",xlab = st_labs, ylab="Cost ($)")
```


```{r}
scatter.smooth(cgdata$ColGrad.,cgdata$EdSpend)

```


```{r}
scatter.smooth(cgdata$ColGrad.,cgdata$LPRFem, xlab = "College Grad %", ylab = "LPRFem", main="LRPFem & College Grad Percentage by State")
text(cgdata$ColGrad.,cgdata$LPRFem,labels = cgdata$State)
```

```{r}
hist(cgdata$LPRFem)
```

```{r}

barplot(cgdata$Metro., main = "Metro Percentage by State",
        xlab = st_labs, ylab = "Percentage")
```


```{r}
barplot(cgdata$Metro., main = "Mean Age of Attendance",xlab = st_labs, ylab = "Age")
```


```{r}
barplot(cgdata$LPRFem, main = "LPRFem by State",xlab = st_labs, ylab = "LPRFem")
```

```{r}
grads = cgdata$ColGrad.
drops = cgdata$Dropout.
spend = cgdata$EdSpend
grads
```


```{r}
drops
```

```{r}
spend
```
Use order() function and set the indices of the percentages of the college graduates in decreasing order. Then take the first ten values for the top ten states with the highest college graduation rates. 
```{r}
order(grads, decreasing = T)
```

```{r}
order(drops, decreasing = T)
```

```{r}
order(spend, decreasing = T)
```
Put these values into vectors
```{r}
# College Graduate Top 10 Percentages by State
# Take the 11 columns' values also
# Indices: 7 21 20 30 6 45 23 29 47 46
cgdata[7,]
cgdata[21,]
cgdata[20,]
cgdata[30,]
cgdata[6,]
cgdata[45,]
cgdata[23,]
cgdata[29,]
cgdata[47,]
cgdata[46,]
```

```{r}
# Top 10 Dropout Percentages by State
# Take the 11 columns' values also
# Indices: 43 17 24 18  5  1 31  4 42 48
cgdata[43,]
cgdata[17,]
cgdata[24,]
cgdata[18,]
cgdata[5,]
cgdata[1,]
cgdata[31,]
cgdata[4,]
cgdata[42,]
cgdata[48,]
```

```{r}
# College Graduate Top 10 Percentages by State
# Take the 11 columns' values also
# Indices: 13  2  7 32 45 30 50 22 23 21
cgdata[13,]
cgdata[2,]
cgdata[7,]
cgdata[32,]
cgdata[45,]
cgdata[30,]
cgdata[50,]
cgdata[22,]
cgdata[23,]
cgdata[21,]

```


```{r}
names(cgdata)
```


Use order() function and to get the indices of the states labeled part of the Northeast region of the United States. Then take the states that are part of this region which have a 1 value and set decreasing to True.
```{r}
order(cgdata$Neast == 1, decreasing = T)
# Indices: 7  8 19 20 21 29 30 32 38 39 45
# Count: 11
```

```{r}

```



Use order() function and to get the indices of the states labeled part of the Southeast region of the United States. Then take the states that are part of this region which have a 1 value and set decreasing to True.
```{r}
order(cgdata$Seast == 1, decreasing = T)
# Indices for Southeast US region: 1  4  9 10 17 18 24 33 40 42 46 48
# Count: 12

```
Use order() function and to get the indices of the states labeled part of the West region of the United States. Then take the states that are part of this region which have a 1 value and set decreasing to True.
```{r}
order(cgdata$West == 1, decreasing = T)
# Indices for West US region: 2  3  5  6 11 12 26 28 31 37 43 44 47 50
# Count 14
```
Use order() function and to get the indices of the states labeled part of the Midwest region of the United States. Then take the states that are part of this region which have a 1 value and set decreasing to True.
```{r}
order(cgdata$Midwest == 1, decreasing = T)
# Indices for Midwest US region: 13 14 15 16 22 23 25 27 34 35 36 41 49
# Count: 13
```


```{r}


```




















